A theorical-practical approach to manifold learning

### 2018-11-05 DimReduction/Mds/Documentation/Article_mds.tex
Adding introduction to Mds documentation

Dimensionality reduction is a technique

\cite{manifold}

\cite{dimension}

\cite{mds}

@@

    \begin{equation*}
            \hspace{-1cm}
        \begin{aligned}
            Tr[(V^tAV)^2] &= Tr[A^2]\\
            Tr[(V^t(V\Lambda V^t - Q \hat{\Lambda} Q^t)V)^2] &= Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]\\
            Tr[(V^tV\Lambda V^tV - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
            Tr[(I\Lambda I - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
        \end{aligned}
    \end{equation*}

@@

    \begin{equation*}
        \begin{aligned}
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n d_{ij}^2
            &=
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n [x_i^t x_i - 2 x_i^t x_j + x_j^t x_j]\\
            &=
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i - \frac{2(0)}{n^2} +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
            &=
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
            &=
            \frac{n}{n^2}\sum_{i=1}^n x_i^t x_i +  \frac{n}{n^2}\sum_{j=1}^n x_j^t x_j\\
            &=
            \frac{1}{n}\sum_{i=1}^n x_i^t x_i +  \frac{1}{n}\sum_{j=1}^n x_j^t x_j\\
            &=
            \frac{2}{n}\sum_{i=1}^n x_i^t x_i\\
        \end{aligned}
    \end{equation*}

@@

    \begin{equation*}
        \begin{aligned}
            x_i^t x_j
            &=
            \frac{1}{2} \left[ \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - d_{ij}^2 + \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n d_{ij}^2 \right]\\
        \end{aligned}
    \end{equation*}

>>>>

Dimensionality reduction is a technique that aims to summarize the information
of a dataset in the most compact way posible in order to facilitate the analysis
procedures of the data. These procedures can also be seen as a process to find
a set of degrees of freedom that econde most of the dataset variability
\cite{proof}.

Some of the approximations in these types of methods assume that the data that
is being analyzed does not exploit all the dimensionality of the space where
is defined. Due to this fact, another description of the goal that the
dimensionality reduction techniques have is related to find a meaningful low
dimensional structure that is hidden in a multidimentional structure.
\cite{dimension}

A clear example of these type of methods are manifold learning techniques. These
approximations search for a way to perform a non linear dimensionality reduction,
which means that they do not assume that the structure of the original dataset
is encoded based on euclidean distance meassuares. Instead these algorithms try
to express their relationships in terms of geodesic distances, which are
distances meassured over the manifold surface. The result of this intend is 
to preserve the intrinsic structure of the data when the reduction is perfomed,
allowing to conserve the implicit information that was hidden in the original
structure of the dataset. \cite{manifold}.

One example of a manifold learning algorithm is called multidimensional
scaling \cite{mds}, this method addresses the problem by trying to find a low
dimensional distance matrix that is as close as possible to a multidimensional
version that encodes the relationship between the points in the original space.
Then the algorithm uses the founded matrix to build a correspondent configuration
of points in a low dimentional space that will act as a less complicated
representation of the original dataset.

This work aims to develop the theorical and practical aspects of the method
described above, with the objective of build a method that can reduce a
multidimensional manifold to a low dimensional representation. To achieve this
in the first part of the work a theorical framework is developed, in order to
understand all the aspects of the method. Then an implementation section will
take place, followed by a result section. Later, some conclusions are
pressented and finally an apendix is added to present some proofs of the
steps that are used in the theorical framework.

@@

    \begin{equation*}
        \begin{aligned}
            Tr[(V^tAV)^2] &= Tr[A^2]\\
            Tr[(V^t(V\Lambda V^t - Q \hat{\Lambda} Q^t)V)^2] &= Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]\\
            Tr[(V^tV\Lambda V^tV - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
            Tr[(I\Lambda I - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
        \end{aligned}
    \end{equation*}

@@

    \begin{equation*}
        \begin{aligned}
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n d_{ij}^2
            &=
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n [x_i^t x_i - 2 x_i^t x_j + x_j^t x_j]\\
            =
            \frac{1}{n^2}&\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i - \frac{2(0)}{n^2} +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
            =
            \frac{1}{n^2}&\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
            =
            \frac{n}{n^2}&\sum_{i=1}^n x_i^t x_i +  \frac{n}{n^2}\sum_{j=1}^n x_j^t x_j\\
            =
            \frac{1}{n}&\sum_{i=1}^n x_i^t x_i +  \frac{1}{n}\sum_{j=1}^n x_j^t x_j\\
            =
            \frac{2}{n}&\sum_{i=1}^n x_i^t x_i\\
        \end{aligned}
    \end{equation*}

@@

    \begin{equation*}
        \hspace{-0.5cm}
        \begin{aligned}
            x_i^t x_j
            &=
            \frac{1}{2} \left[ \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - d_{ij}^2 + \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n d_{ij}^2 \right]\\
        \end{aligned}
    \end{equation*}
